Quantum Mechanics Assignment Three

### 2016-02-02 quantum/latex/a4.tex
almost done!

>>>>

\section{10.24: Degenerate Perturbation}
Here we are given just the perturbed Hamiltonian of a system and are asked to find the first
nonvanishing energy correction to each of the four states. The perturbed Hamiltonian is:

\begin{equation*}
  H \doteq
  \begin{pmatrix}
    3 & \epsilon & 0 & 0\\
    \epsilon & 3 & 2\epsilon & 0\\
    0 & 2\epsilon & 5 & \epsilon\\
    0 & 0 & \epsilon & 7\\
  \end{pmatrix}
\end{equation*}

\textbf{a.}
We are told that there are no diagonal perturbation elements, and so because the Hamiltonian is in
the original basis we write our states as follows:

\begin{equation*}
  3V_0,
  \begin{pmatrix}
    1 \\ 0 \\ 0 \\ 0\\
  \end{pmatrix}
  \hspace{2cm}
  3V_0,
  \begin{pmatrix}
    0 \\ 1 \\ 0 \\ 0\\
  \end{pmatrix}
  \hspace{2cm}
  5V_0,
  \begin{pmatrix}
    0 \\ 0 \\ 1 \\ 0\\
  \end{pmatrix}
  \hspace{2cm}
  7V_0,
  \begin{pmatrix}
    0 \\ 0 \\ 0 \\ 1\\
  \end{pmatrix}
  \hspace{2cm}
\end{equation*}

\textbf{b.}
First-order corrections always manifest as terms on the Hamiltonian's diagonal. Our diagonal has
no such terms, and so all corrections are 2nd-order or higher. We first solve for the nondegenerate
$\ket{3}$ and $\ket{4}$ state energy corrections:

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
  \frac{|H'_{3,2}|^2}{E_2 - E_3} +
  \frac{|H'_{3,4}|^2}{E_4 - E_3}\\
  &= \frac{4\epsilon^2}{-2V_0} + \frac{\epsilon^2}{-4V_0}\\
  &= \frac{-9\epsilon^2}{4V_0}
\end{align*}

\begin{align*}
  E_4^{(2)} &=
  \frac{|H'_{4,1}|^2}{E_1 - E_4} +
  \frac{|H'_{4,2}|^2}{E_2 - E_4} +
  \frac{|H'_{4,3}|^2}{E_3 - E_4}\\
  &= \frac{-\epsilon^2}{2V_0}
\end{align*}
